# Housing Prices Submission
***

## 1. Importing the needed libraries
```{r eval=FALSE, Packages}
library(dplyr)
library(ggplot2) # Data visualization
library(VIM)  # Plots missing data
library(mice)
library(ggthemes)
library(scales) # Scaling yaxis
library(caret)
```

## 2. Importing data
```{r}
train <- read.csv('/Users/wafic/Downloads/data/House_Prices/train.csv', stringsAsFactors = F, na.strings=c("","NA"))
test <- read.csv('/Users/wafic/Downloads/data/House_Prices/test.csv', stringsAsFactors = F, na.strings=c("","NA"))
submission <- read.csv('/Users/wafic/Downloads/data/House_Prices/sample_submission.csv', stringsAsFactors = F, na.strings=c("","NA"))
```

```{r}
str(train)
```

## 3. Investigating dataset
```{r}
nulls <- function(x){
  sum(is.na(x))
}

sapply(train, nulls)
```

```{r}
aggr(train, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(train), cex.axis=.6, gap=4, ylab=c("Histogram of missing data","Pattern"))
```


## 4. Imputing missing datapoints
```{r}
train$Alley[is.na(train$Alley)] <- 'None'

train$FireplaceQu[is.na(train$FireplaceQu)] <- 'None'

train$Fence[is.na(train$Fence)] <- 'None'

train$MiscFeature[is.na(train$MiscFeature)] <- 'None'

train$PoolQC[is.na(train$PoolQC)] <- 'None'

train$GarageCond[is.na(train$GarageCond)] <- 'None'
train$GarageQual[is.na(train$GarageQual)] <- 'None'
train$GarageFinish[is.na(train$GarageFinish)] <- 'None'
train$GarageType[is.na(train$GarageType)] <- 'None'
train$GarageYrBlt[is.na(train$GarageYrBlt)] <- 0

train$BsmtFinType1[is.na(train$BsmtFinType1)] <- 'None'
train$BsmtFinType1[is.na(train$BsmtFinType1)] <- 'None'
train$BsmtFinType2[is.na(train$BsmtFinType2)] <- 'None'
train$BsmtExposure[is.na(train$BsmtExposure)] <- 'None'
train$BsmtCond[is.na(train$BsmtCond)] <- 'None'
train$BsmtQual[is.na(train$BsmtQual)] <- 'None'

train$MasVnrType[is.na(train$MasVnrType)] <- 'None'
```

```{r}
pmiss <- function(x){
  sum(is.na(x))/length(x)*100
}
pmiss(train$LotFrontage)
```

```{r}
tempdata <- mice(train, m=5, maxit=50, meth='pmm',seed=500)
tempdata$imp$LotFrontage
```


```{r}
completedData <- complete(tempdata,1)
train$LotFrontage <- completedData$LotFrontage
train$MasVnrArea <- completedData$MasVnrArea

train$Electrical <-factor(train$Electrical)
table(train$Electrical)
```

```{r}
train$Electrical[is.na(train$Electrical)] <- 'SBrkr'
```


```{r}
sapply(train, nulls)
```

```{r}
aggr(train, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(train), cex.axis=.6, gap=4, ylab=c("Histogram of missing data","Pattern"))
```

```{r}
str(train)
```

```{r}
ggplot(train, aes(x='', y=SalePrice))+
  geom_boxplot()+
  scale_y_continuous('House Prices',labels = dollar)+
  theme_minimal()
```


```{r}
ggplot(train, aes(x=factor(SaleCondition), y=SalePrice))+
  geom_boxplot()+
  scale_y_continuous('House Prices',labels = dollar)+
  theme_minimal()
```

```{r}
ggplot(train, aes(x=factor(SaleType), y=SalePrice))+
  geom_boxplot()+
  scale_y_continuous('House Prices',labels = dollar)+
  theme_minimal()
```

I can see 4 types of variable:
1. Neigjborhood features
2. House features
3. Sales features

```{r}
train[sapply(train, is.numeric)]
```

```{r}
train[sapply(train, is.character)]
```

1. CONVERT CHARACTER TO FACTOR IF REQURIED
2. GROUP VARIABLE INTO GROUPS AS BELOW

I can see 4 types of variable:
1. Neigjborhood features
2. House features
3. Sales features