# Predict Future Sales Submission
***
The task is to forecast the total amount of products sold in every shop for the 
test set. Note that the listof shops and products slightly changes every month. 
Creating a robust model that can handle such situations is part of the challenge

You need to forecast the sales for these shops and products for November 2015

```{r}
library(dplyr)
library(ggplot2) # Data visualization
library(forecast)
```

```{r}
item_cat <- read.csv('/Users/wafic/Downloads/Predict_Sales_Kaggle/item_categories.csv', stringsAsFactors = FALSE)
items <- read.csv('/Users/wafic/Downloads/Predict_Sales_Kaggle/items.csv', stringsAsFactors = FALSE)
test <- read.csv('/Users/wafic/Downloads/Predict_Sales_Kaggle/test.csv', stringsAsFactors = FALSE)
train <- read.csv('/Users/wafic/Downloads/Predict_Sales_Kaggle/sales_train.csv', stringsAsFactors = FALSE)
shops <- read.csv('/Users/wafic/Downloads/Predict_Sales_Kaggle/shops.csv', stringsAsFactors = FALSE)
```

```{r}
str(train)
```


## Exploring the data
```{r}
dim(item_cat);dim(items)
```
We have 22,170 items categorized into 84 categories

#### Merging column id with train data
```{r}
head(items)
```

```{r}
train_data <- merge(train, items, by = "item_id")
```
Added category column to the train data

```{r}
str(train_data)
```

```{r}
train_data$day <- format(as.Date(train_data$date,format="%d.%m.%Y"), "%d")
train_data$month <- format(as.Date(train_data$date,format="%d.%m.%Y"), "%m")
train_data$year <- format(as.Date(train_data$date,format="%d.%m.%Y"), "%Y")

train_data$day <- factor(train_data$day)
train_data$month <- factor(train_data$month)
train_data$year <- factor(train_data$year)

train_data$shop_id <- factor(train_data$shop_id)
train_data$item_id <- factor(train_data$item_id)
train_data$item_category_id <- factor(train_data$item_category_id)

str(train_data)
```


```{r}
head(train_data)
```

```{r}
yearly <- group_by(train_data, year)
summarise(yearly, items_sold = sum(item_cnt_day))
```

```{r}
str(test)
```

```{r}
head(test, 50)
```


